  ---
title: "R Final classProject Code"
author: "Ruth Holley"
date: "6/1/2019"
output: html_document
---
Load our data Part I
Cancer Survival Rates in the United States
```{r}
library(dplyr)
library(ggplot2)
#Survival Rates in the United States
cancersurvival <- read.csv("/home/ruth/Downloads/cancer-survival-Rates-in-usa.csv",header = TRUE, sep = ',')

summary(cancersurvival)
str(cancersurvival)
dim(cancersurvival)
```

Load our data Part II
Share of population suffering from cancer types
```{r}
cancer<- read.csv("/home/ruth/Downloads/share-of-population-suffering-from-cancer-types.csv",header = TRUE, sep = ',')
#This is a worldwide population dataset. It has been filtered below to only included United States
summary(cancer)
str(cancer)
dim(cancer)
```

1. Which are the top five cancers that are more prevalent in the United States#Question not used but will reseach for my own purpose later***
```{r }
c1<-cancer%>%
  filter(Entity =="United States")%>%
group_by("Year")
c1
summary(c1)
```

```{r}
ggplot(c1, aes(x =Year, y=Breast.cancer)) +
geom_bar(stat = "identity")
```


1. Which cancers have the greatest survival rate in the United States

```{r}
c2<-cancersurvival%>%
  select(Survival.Rate,Year,CancerType)%>%
  group_by(Year,CancerType)%>%
  arrange(desc(Survival.Rate))
c2
	
```


```{r}
ggplot(c2, aes(x =CancerType, y=Survival.Rate,fill =CancerType,)) +
geom_bar(stat = "identity")  + ggtitle("Which cancers have the greatest survival rate in USA")
```

2. Which types of cancer affects mostly women versus men?
```{r }

'%!in%' <- function(x,y)!('%in%'(x,y))

c3<-cancersurvival%>%
  select(Survival.Rate,Year,Gender,CancerType)%>%
  filter(Gender %in% c("females","males") & CancerType %!in% "All cancers" )
  
c3

```

```{r}

ggplot(c3, aes(x=Gender, y=CancerType, fill=Survival.Rate)) +
geom_bar(stat="identity", position=position_dodge()) +
    ggtitle("Which types of cancer affects mostly women versus men?")




```

3. Which type of cancer has the highest survival rate in both genders in 10 year intervals.

women**Don't use**
1973,1983,1995,2006.2013
```{r}
cancersurvival%>%
  select(Survival.Rate,Year,Gender,CancerType)%>%
  filter(Gender == "females" & Year==2013)%>%
  group_by(CancerType)%>%
  arrange(desc(Survival.Rate))
```



**Graphs starts here for 10 year intervals** 
Top 5 cancer rates-females
```{r}
t5<-cancersurvival%>%
  select(Survival.Rate,Year,Gender,CancerType)%>%
  filter(Year %in% c(1973, 1983), CancerType %in% c("Thyroid","Skin","Breast","Bladder","Cervix uteri"), Gender=="females")
t5
```

Female-Top 5 cancers year 1973-1983--GGPLOT
```{r}
ggplot(t5, aes(x =Year, y=Survival.Rate,fill =CancerType,)) +
geom_bar(stat = "identity") + ggtitle("Female Cancer Survial Rate from 1973-1983?")
```

Top 5 cancer rates-males
```{r}
t6<-cancersurvival%>%
  select(Survival.Rate,Year,Gender,CancerType)%>%
  filter(Year %in% c(1973, 1983), CancerType %in% c("Thyroid","Skin","Prostate","Bladder","Myeloma"), Gender=="males")
t6
```

Top 5 cancer-males-GGPLOT
```{r}
ggplot(t6, aes(x =Year, y=Survival.Rate,fill =CancerType)) +
geom_bar(stat = "identity") + ggtitle("Male Cancer Survial Rate from 1973-1983?")
```

Top 5 Cancer Survival Rates--1995-2006 Female graph code
```{r}
t7<-cancersurvival%>%
  select(Survival.Rate,Year,Gender,CancerType)%>%
  filter(Year %in% c(1995, 2006), CancerType %in% c("Thyroid","Skin","Breast","Bladder","Cervix uteri"), Gender=="females")
t7
```

1995-2006--Female-GGPLOT
```{r}
ggplot(t7, aes(x =Year, y=Survival.Rate,fill =CancerType)) +
geom_bar(stat = "identity")  + ggtitle("Female Cancer Survial Rate from 1995-2006?")
```

Top 5 Cancer Rates-1995-2006 male graph code
```{r}
t8<-cancersurvival%>%
  select(Survival.Rate,Year,Gender,CancerType)%>%
  filter(Year %in% c(1995, 2006), CancerType %in% c("Thyroid","Skin","Prostate","Bladder","Colon and rectum"), Gender=="males")
t8
```

1995-2006 male graph GGPLOT
```{r}
ggplot(t8, aes(x =Year, y=Survival.Rate,fill =CancerType)) +
geom_bar(stat = "identity")  + ggtitle("Male Cancer Survial Rate from 1995-2006?")
```

Top 5 cancer survival rates-females (2006-2013)
```{r}
t9<-cancersurvival%>%
  select(Survival.Rate,Year,Gender,CancerType)%>%
  filter(Year %in% c(2006,2013), CancerType %in% c("Thyroid","Skin","Breast","Bladder","Cervix uteri"), Gender=="females")
t9
```

2006-2013-Female-GGPLOT
```{r}
ggplot(t9, aes(x =Year, y=Survival.Rate,fill =CancerType)) +
geom_bar(stat = "identity") + ggtitle("Female Cancer Survial Rate from 2006-2013?")
```

To 5 Cancer Survival Rates 2006-2013 male graph code
```{r}
t10<-cancersurvival%>%
  select(Survival.Rate,Year,Gender,CancerType)%>%
  filter(Year %in% c(2006, 2013), CancerType %in% c("Thyroid","Skin","Prostate","Bladder","Colon and rectum"), Gender=="males")
t10
```

2006-2013-male-GGPLOT
```{r}
ggplot(t10, aes(x =Year, y=Survival.Rate,fill =CancerType)) +
geom_bar(stat = "identity")+ ggtitle("Male Cancer Survial Rate from 2006-2013?")

```











histogram ggplot
```{r}
ggplot(cancersurvival, aes(x =Year, y=Survival.Rate,fill =CancerType)) +
geom_bar(stat = "identity")
```



#mem
```{r}
cancersurvival%>%
  select(Survival.Rate,Year,Gender,CancerType)%>%
  filter(Gender == "males" & Year==2013)%>%
  arrange(desc(Survival.Rate))
```

4. Has the survival rate increased over a mostly 50 years interval?
```{r}
t11<-cancersurvival%>%
  select(Survival.Rate,Year,Gender,CancerType)%>%
  filter(Year %in% c(1963, 2013)) 

t11
```


```{r}
ggplot(t11, aes(x =Year, y=Survival.Rate, fill=Survival.Rate)) +
geom_bar(stat = "identity")+
  ggtitle("Has the survival rate increased over a 50 year interval?")
```


5. Is race a factor in cancer survival rate?
```{r}
t12<-cancersurvival%>%
  select(Survival.Rate,Year,Race,CancerType)%>%
  filter(Race %in% c("Black","White"))%>%
  group_by(CancerType)%>%
  arrange(desc(Survival.Rate))
t12
```


```{r}
ggplot(t12, aes(x =Race, y=Survival.Rate, fill=Race)) +
geom_bar(stat = "identity") +ggtitle("Could Race be a factor in Cancer Survival Rates?")
    
```








